---
title: "05_QuantLing_Section"
author: "Eric Wilbanks"
date: "2/21/2020 - 2/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{css echo=FALSE}
.example{
  background-color: rgba(30,206,124,.5);
}
```

## 1. Goals for Section
  1. Reminder: HW Submission Format as RMarkdown
  1. Reviewing Linear Models
  1. Visualing Model Fit
  1. Homework Workshop

## 2. Linear Model Review
Let's consider the `ratings` dataset from `languageR`, in which people rated the subjective size of different nouns.

```{r}
library(lme4)
library(languageR)
library(ggplot2)
data(ratings)
head(ratings)
```


```
1. Are ratings of size significantly predicted by how familiar objects are and what class they are in? Fit a multiple linear model to answer this question.
2. What type of variable (categorical or continuous) are each of these terms?
3. How do we get the t-values in the column?
4. What are the significant differences?
5. What parts of this output do you have questions about?
```

```{r}
ggplot(ratings, aes(meanFamiliarity,meanSizeRating,color=Class)) + geom_point() + geom_smooth(method='lm')
```

```
1. Compare the plot above to the model output. What does our model intercept correspond to here?
2. Compare the coefficients from the model output; how do they map onto the plot?
```

## 3. Visualizing Model Fit

```
When considering how well our model fits the data, what should we look at or test?
```

Let's plot the model from earlier and discuss whether it matches these criteria.
```{r}
#plot(modelName)
```

```{r}
# we can also add the fitted values to the original data-frame to explore further in plots
#ratings$fitted <- fitted(modelName)
```

```
1. Create a new factor in the ratings df that is the residuals from the model.
2. Create a ggplot which compares the fitted and actual values for meanSizeRatings (consider using `geom_abline(slope=1)` )
3. Create a plot that compares the residuals between the two Classes. Is the model better at predicting one class than another?
4. Create a plot that compares the residuals between different levels of Familiarity. Is the model better at predicting objects of certain levels of familiarity?
```
## 4. Homework Workshop

