---
title: "01_QuantLing_Section"
author: "Eric Wilbanks"
date: "1/24/2020 - 1/27/2020"
output: 
  html_document: 
    theme: paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{css echo=FALSE}
.example{
  background-color: rgba(30,206,124,.5);
}
```
## 1. Goals for Section

- understand purpose of `R` and `RStudio`
- troubleshoot/confirm installation of `R` and `RStudio`
- (briefly) discuss alternatives such as base R, Jupyter
- explore basics of R syntax, variables, functions, and packages

## 2. R and RStudio Purpose

### R
`R` is a language and platform for doing statistical analysis and data visualization. People like using `R` because

1. It is free and open source
2. It is well-supported and broadly used
3. It easily allows for extensions (you could create your own packages to do a specific task!)
4. It has a shallow learning curve

```
Why not use Python and Pandas for everything?
```

We might see a shift towards doing everything in Python in the future; right now, a lot of statistical methods are only formulated in R; there's a long history of development and error-checking code implementation in R. 

### RStudio

RStudio is an alternative GUI (graphical user interface) for R, and provides user-friendly features that are not found in basic R interpreters. 

Let's discuss some of those features now.

- Console
- Environment tab
- Plots
- Organize and resize windows you need

## 3. Numeric Operations
The most basic use of R is for numeric calculations.

```{r}
10+5
12/4
3*2
5^2
(10/2)-1
```


## 4. Objects, Data-Structures, Variable Assignment

R is an "object-oriented" language, meaning we define objects that contain data, and can refer to and manipulate these objects by their name.

There are **various data-types**; we'll mainly focus on the first two: 

- character
- numeric
- logical
- integer
- complex

Note the differences between the `cat` variable, which we've filled with numeric data and the `dog` variable, which we've filled with character data. 

```{r}
cat <- 8
dog <- "8"
```

```{r}
cat
```

```{r}
dog
```

```{r, eval = FALSE}
cat + 8
dog + 8
```

There are **various data-structures** that we can fill with data.

- Vectors (we'll focus on these for now)
- Data frames
- Matrices
- Lists

```{r}
trees <- c("elm", "oak", "cedar", "eucalyptus")
trees
```

Vectors are ordered collections of elements. The `trees` vectors has 4 elements, each a character  type data-point. 

We can reference specific elements of a vector with an index.

```{r}
trees[4]
```

**Note that unlike Python, R is 1-index. The first element is indexed with [1]**

## 5. Functions
Functions in R are handy wrappers for complicated tasks. The standard format is lower-case text followed by parentheses filled with arguments. 

Let's consider the `rnorm` function, which randomly samples from a normal distribution (gaussian). 
```{r}
rnorm(3)
```

If you encounter a function and want to understand what it does, enter a `?` before the function name in the command line to open the documentation.

**This is sometimes less helpful than a google search**

```{r, eval=FALSE}
?rnorm
```

1. Use the `rnorm` documentation to find out what the default parameters of the normal distribution it's drawing from are. 
2. Then, use `rnorm` to generate 6 random samples from a Gaussian with mean of 50 and standard deviation equal to 0.4

## 6. Packages

Packages are collections of functions, objects, and other code which are not included in the "basic" set of R commands.

Packages are one of two types:

1. Included with the default installation, but must be loaded
1. Not included with the default installation, must be installed before they can be loaded.

To load a package, use the `library()` function. Whenever you restart R, you'll have to reload any packages you want to use that session.

```{r, eval=FALSE}
library(lme4)
```

```
What does the lme4 package do?
```

If you don't have a package installed, you can install it with the `install.packages()` function. **Note the quotation marks!** 

You only have to install a package once.

```{r, eval=FALSE}
install.packages('ggplot2')
```

