---
title: "10_QuantLing_Section"
author: "Eric Wilbanks"
date: "4/10/2020 - 4/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{css echo=FALSE}
.example{
  background-color: rgba(30,206,124,.5);
}
```

## 1. Goals for Section

1. Regular Expression Review
1. Regular Expressions in R
1. Capture Groups and Regex Search and Replace

## 2. Regex Review

### 2.1 Special Characters

**Quantifiers**

1. a* (0 or more of previous character/group, in this case `a`)
1. a+ (1 or more of previous character/group, in this case `a`)
1. a? (0 or 1 of previous character/group, in this case `a`)
1. a\{n\} (n occurences of previous character/group, in this case `a`)
1. a\{n,\} (n or more occurences of previous character/group, in this case `a`)
1. a\{m,n\} (between m to n occurences (inclusive) of previous character/group, in this case `a`)

**Position Within String: "Anchors"**

1. ^ (start of string)
1. $ (end of string)
1. \\b (word boundary)
1. \\n (end of line character)
1. \\r (carriage return character)

**Grouping**

1. (a|b|c) (any one character that is a, b, or c; parentheses are **capture groups** [more later])
1. [abc] (any one character that is a, b, or c; brackets are **character classes**)
1. [^abc] (any one character that isn't a, b, or c)


**Common Symbols**

1. . (any single character, including space!)
1. \\ (escape character)
    1. \\$ (just literal dollar sign)
    1. \\. (just literal period)
    
### 2.2 Regex Testers

A great resource for learning is using online regex compilers/testers, like https://regex101.com/. 

Different programming languages may have slightly different flavors or resolutions for more complicated regular expressions, but they should all be the same for the common characters we're using.

Let's try out the following examples as a group:

1. https://regex101.com/r/DiYV4x/3
1. https://regex101.com/r/wFipxo/1
1. https://regex101.com/r/eeHtOL/1

### 2.3 Regex in R

```{r}
# base function for doing pattern testing is grep()
# syntax is: grep('pattern', 'test vector/strings')

################
# default output is which string/element matched
################
grep('^a', 'angel')

################
# here, strings 1 and 4 match the pattern
################
grep('^a', c('alphabet','cat','Animal','applies'))

################
# set value = T to get value, not index of matches
################
grep('^a', c('alphabet','cat','Animal','applies'), value = T)
```

```{r, eval=F}
################
# On your own, come up with regular expressions that will match the correct elements
################

################
# define pattern
r = ''
  
# match these
grep(r, c('Tarah','tarah','Tara','tara','hey tara'))

# don't match these
grep(r, c('Haphtarahs','tarantula','xtara'))
################
```

```{r, class.source='example'}
# define pattern
r = '(^| )[Tt]arah?$'
  
# match these
grep(r, c('Tarah','tarah','Tara','tara','hey tara'))

# don't match these
grep(r, c('Haphtarahs','tarantula','xtara'))


#############
# importantly, grep() doesn't treat start/end of string as word boundaries, 
# so '\b[Tt]arah?\b' won't work
#############
```


## 3. Capture Groups and Regex Search and Replace

One of the most powerful uses of regular expressions is Search and Replace.

Imagine if you wanted to find all instances of days of the week in a corpus and replace them with capitalized versions of themselves.

1. monday $\rightarrow$ MONDAY
1. Monday $\rightarrow$ MONDAY
2. Friday $\rightarrow$ FRIDAY

You could do a bunch of search and replaces manually, but this would be exhausting and error-prone. Instead, let's rely on the power of regular expressions to be more efficient.

### 3.1 Capture Groups

So far, we've just been asking you all to specify what lines or strings contain a pattern we want. But lines also contain material besides our pattern and match.

```{r}
# what part of the string matches the pattern?
grep('[4-9]+Hz',c('a total of 45Hz change'), value = T)

#############
# the substring '45Hz' matches our pattern
#############

# let's explicitly pull out the matching parts using the following new functions:
x <- 'a total of 45Hz change'
regmatches(x, regexpr('[4-9]+Hz', x))
```

But what if I don't care about the Hz marks at all? I need them in my pattern to avoid getting any old numbers, but I want to separate out just the numbers. 

We can use the `sub` function and capture groups to accomplish this!

```{r}
sub('([4-9]+)Hz', '\\1', x)
```

```
1. How is the regex pattern in the sub() case different from the pattern in the two previous cases?
2. What do you think '\\1' corresponds to?
```

```{r, class.source='example'}
#1) we have parentheses (a capture group) around the number part of the pattern
#2) \\1 corresponds to the first capture group 
# [in r we have to escape the backslash in this contexts; \1 is the typical reference to the first capture group]
```

```{r, eval=T}
##############
# What will this output be?
##############
sub('([4-9]+)(Hz)', '\\2', x)
```

```{r, eval=T}
##############
# What will this output be?
##############
sub('([4-9]+)(Hz)', '\\2\\1\\1', x)
```

```{r}
##############
# our replacement doesn't just have to contain references to capture groups
# we could do anything!
##############
sub('([4-9]+)(Hz)', 'about 3 avocados are \\1(ish) \\2 tall without any', x)
```

### 3.2 Extra Practice - Corpus Regex

The `moby` object is a vector where each element is a word from the novel Moby Dick. 

```{r, eval=F}
library(languageR)
data(moby)

# First, what does this regex do?
matches <- grep('[^ ]day$', moby, value=T)
matches

# Second, find a way to replace all of the "day" substrings in `matches` with "DAY"
sub('','', matches)

# Finally, find a regular expression that will only match days of the week in `matches`, and exclude all of the others

matches2 <- grep('', matches, value =T)
matches2
```

```{r, class.source='example'}
library(languageR)
data(moby)

# this regex matches all of the elements of this vector where we have a character that's not a space before 'day', which is at the end of the string
matches <- grep('[^ ]day$', moby, value=T)
matches

#replace all of the "day" substrings in `matches` with "DAY"
sub('day','DAY', matches)



#find a regular expression that will only match days of the week in `matches`, and exclude all of the others

matches2 <- grep('^[SM][a-z]+day$', matches, value =T)
matches2

# this regex gets the job done for matches2, but is prone to false positives for other datasets
# for example, it would match 'Someday'
# we'd probably want the following instead:

matches3 <- grep('^(Sun|Mon|Satur)day$', matches, value =T)
matches3 
```