---
title: "03_QuantLing_Section"
author: "Eric Wilbanks"
date: "2/07/2020 - 2/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{css echo=FALSE}
.example{
  background-color: rgba(30,206,124,.5);
}
```
## 1. Goals for Section

- remaining data frame points
- loading/saving data
- visualization with ggplot2

## 2. Data Frame Errata

Another useful operator is the `%in%` operator. This can be more efficient that many chained `or` statements.

```{r}
library(languageR)
data(sizeRatings)
fruits <- sizeRatings[sizeRatings$Word %in% c('cherry','lemon','pear','blackberry'),]
```

We can also create new factors! In the code below, we add a new factor which has a value of `familiar` if `sizeRatings$MeanFamiliarity` is greater than 0.7, and `unfamiliar` if not. 

```{r}
sizeRatings$binary_MF <- ifelse(sizeRatings$MeanFamiliarity > 0.7, 'familiar', 'unfamiliar')
```

```
What differences are there between $Language and $binary_MF when you call str(sizeRatings) ?
```
```{r class.source="example", eval=FALSE}
# sizeRatings$binary_MF is a character vector, while sizeRatings$Language is a factor.
# this is because ifelse() outputs a character vector; we can convert it using the following:
sizeRatings$binary_MF <- as.factor(sizeRatings$binary_MF)
```

## 3. Loading and saving dataframes

A common format for data is `.csv`, ("comma separated values"). 

```
Download the `PresuppositionType_data.csv` file from bCourses. 

What does this file look like if you open it in a text editor?
```

### 3.1 Reading in Data Frame
```{r, eval = FALSE}
presup <- read.csv(file.choose())
```

You can also specify the path to the file if you have that information. This is always specified relative to your working path.

```
What is a working path?
```
```{r class.source="example", eval=FALSE}
# The current directory or folder that you're working in
```

```{r}
presup <- read.csv('PresuppositionType_data.csv')
```

You can query your current working directory if you're not sure, and can set it to a specific directory. You can also permanently change your working directory in RStudio using **Tools > Global Options**

```{r}
getwd()
#setwd('/home/eric/TotallyRealFolder')
```

### 3.2 Saving a data frame (as csv)
```{r, eval = FALSE}
sub <- presup[1:20,]
write.csv(sub, file = 'new_df.csv',row.names=F)
write.csv(sub, file = 'new_df_test.csv', row.names=T) # default option for row.names; what's changed?
```
```{r class.source="example", eval=FALSE}
# The default behavior is to save row.names (numbers) as a column in the csv file. 
# This can be an issue because the default for read.csv() is that there aren't row.names.
# It's a good idea to always write.csv() with row.names = F
```

## 4. Visualization with ggplot2
There are various ways to make visualizations in R. We'll be using a powerful and well documented package, `ggplot2`.

```{r}
# install the package (if you haven't ever done this)
#install.packages('ggplot2')

# load the packages we'll need
library(ggplot2)

# load the Dutch ge- data
data(durationsGe)
```

```{r}
# simple scatter-plot
ggplot(durationsGe, aes(DurationOfPrefix,SpeechRate)) + geom_point()
```

The first part of the command `ggplot(durationsGe, aes(DurationOfPrefix,SpeechRate))` establishes the data that's being used and which factors are mapped to the x and y dimensions.

The second part `+ geom_point()` adds a geom to the graph. You can have multiple geoms that layer on the same plot. 

```
Trying adding geom_smooth(method='lm') to the plot you just made.
```
```{r class.source="example"}
ggplot(durationsGe, aes(DurationOfPrefix,SpeechRate)) + geom_point() + geom_smooth(method='lm')
```

You can change the labels of various elements by creating a label layer, using `labs()` 

```{r}
ggplot(durationsGe, aes(DurationOfPrefix,SpeechRate)) + geom_point() + labs(x='Duration of Ge- prefix', y='Syllables per Second', title='Duration of Ge- by Speech Rate')
```

Let's try a basic density plot

```{r}
ggplot(durationsGe, aes(DurationOfPrefix)) + geom_density()
```

We can also use other factors that aren't x or y dimensions to break down the data:

```{r}
ggplot(durationsGe, aes(DurationOfPrefix, color=Sex)) + geom_density()
```

```
What happens when you replace `color` with `lty`?, `fill`?
```
```{r class.source="example", eval = FALSE}
# color = the line color
# lty = line type
# fill = fill color
```

```{r}
# basic box-plot
ggplot(durationsGe, aes(Sex,DurationOfPrefix, fill=Sex)) + geom_boxplot(alpha=0.5)
```

```
What is the third box-plot?

Find which row(s) of the data-frame are responsible for this third plot. Create a sub-set of the data-frame that avoids this issue, and recreate the plot.
```
```{r class.source="example"}
# there is only one row that has a value of Sex of `NA`.
sub <- durationsGe[durationsGe$Sex %in% c('female','male'),]
ggplot(sub, aes(Sex,DurationOfPrefix, fill=Sex)) + geom_boxplot(alpha=0.5)
```

```{r}
# basic histogram
ggplot(durationsGe, aes(YearOfBirth,fill=Sex)) + geom_histogram()
```

```
What does this plot correspond to?

What happens when we add facet_grid(~Sex) to the plot? What if we add facet_grid(Sex~.) instead?

Create an equivalent density plot and consider how it compares to the histogram.
```

```{r class.source="example"}
# corresponds to a count of the number of rows that have a value of birthyear equal to the year; split up by sex of speaker

ggplot(sub, aes(YearOfBirth,fill=Sex)) + geom_histogram() + facet_grid(~Sex)
ggplot(sub, aes(YearOfBirth,fill=Sex)) + geom_histogram() + facet_grid(Sex~.)
ggplot(sub, aes(YearOfBirth,fill=Sex)) + geom_density(alpha=0.5)


```

## 4. Plot Extensions

```
Let's begin by creating the following plots using the durationsGe dataframe:

1. A scatterplot which shows the relationship between year of birth and speech rate, with a facet by speaker sex.
2. A boxplot of Ge- duration, with an x-axis and fill of NumberSegmentsOnset and custom axis and title labels.
3. A scatterplot of Ge- duration by log() transformed Ge-duration
4. Create a new factor of binary Frequency. ('low' if durationsGe$Frequency is less than the mean of this factor, 'high' if else.). Using this new factor, create a scatterplot of speech rate by ge- duration, faceted by binary frequency. 
```

```{r class.source="example"}
ggplot(sub, aes(YearOfBirth,SpeechRate)) + geom_point() + facet_grid(~Sex) + geom_smooth(method='lm') + labs(x='Birthyear',y='Local Speech Rate\n(Syllables per Second)')

summary(sub)
ggplot(sub, aes(as.factor(NumberSegmentsOnset),DurationOfPrefix,fill=as.factor(NumberSegmentsOnset))) + geom_boxplot() + labs(x='My x-axis',y='y axis custom',title='this is my plot',fill='new legend based on fill')

ggplot(sub, aes(DurationOfPrefix,log(DurationOfPrefix))) + geom_point()

sub$binWF <- as.factor(ifelse(sub$Frequency < mean(sub$Frequency), "low","high"))

ggplot(sub, aes(DurationOfPrefix,SpeechRate)) + geom_point() + facet_grid(~binWF)
```
