---
title: "02_QuantLing_Section"
author: "Eric Wilbanks"
date: "1/31/2020 - 2/03/2020"
output: 
  html_document: 
    theme: paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{css echo=FALSE}
.example{
  background-color: rgba(30,206,124,.5);
}
```

## 1. Goals for Section

- brief introductions and syllabus
- dataframes 
- ~~visualization basics~~

## 2. Introductions and Syllabus

## 3. Dataframes
### 3.1 Basic Indexing
```{r, eval=FALSE}
# Install the languageR package if you haven't yet. 
# Note that we're using # to comment out commands we don't want to execute.

#install.packages('languageR')
```

```{r}
# load the languageR package
library(languageR)
# load the dataset sizeRatings from the languageR package
data(sizeRatings)
```
`sizeRatings` is a dataframe, a data structure in R that is a collection of row and column vectors.
```{r, eval=FALSE}
head(sizeRatings)
str(sizeRatings)
names(sizeRatings)
dim(sizeRatings)
?sizeRatings
```
```
Use the commands above to explore the structure of the dataset. 
1. What does each command do? 
2. What are the relevant rows and columns of the dataset?
```
```{r class.source="example"}
  # head() gives the first 6 rows of the df
  # str() gives a summary of the internal structure of the object
  # names() gives a vector of the column names
  # dim() gives the dimensions of the df
  # ? gives us information about the source of the published dataset
```
We can index elements (cells) and ranges of data.frames in different ways. The first method is similar to how we indexed vectors: `df[rows,columns]`
```{r}
sizeRatings[1,1]
sizeRatings[2,]
sizeRatings[8:10,3]
sizeRatings[4:6,"Word"]
```
Another way to interact with data.frames is to refer to the column vectors (**factors**) by name, using the `$`: `df$factor`
```{r}
head(sizeRatings$Word)
```
```
How would you complete the following commands to make the two subsets equal?
```
```{r, eval=FALSE}
sizeRatings$Word[3:6] == sizeRatings[?,?]
```
```{r class.source="example", eval=FALSE}
sizeRatings$Word[3:6] == sizeRatings[3:6,"Word"]
# OR
sizeRatings$Word[3:6] == sizeRatings[3:6,3]
```

### 3.2 Conditional Indexing
We've been indexing parts of the dataframe by referring to row or column numbers or names. 

We can extend this by indexing parts of the dataframe that meet some condition. For example, I could create a new dataframe that is all of the rows of the sizeRatings where the Word is "pigeon".

```{r}
pigeon <- sizeRatings[sizeRatings$Word == "pigeon",]
```
```{r, eval=FALSE}
#What's happening here?
str(pigeon$Word)
summary(pigeon$Word)
```
Or we could create a subset where all of the size ratings are greater than 6.

```{r}
min(sizeRatings$Rating)
min(sizeRatings[sizeRatings$Rating > 6,]$Rating)
```

Find the `Logical Operators` section on the website https://www.statmethods.net/management/operators.html

Then, create different subsets of the dataframe that for the following criteria.
```
1. A data-frame where the word is either 'blackberry' or 'pigeon'
2. A data-frame where the word is 'ant' and the participants are non-Native speakers of English.
3. Calculate the mean of familiarity ratings for all plant class items.
4. Calculate the mean of familiarity ratings for all animal class items.
```
```{r class.source="example", eval=FALSE}
df1 <- sizeRatings[sizeRatings$Word == 'blackberry' | sizeRatings$Word == 'pigeon',]

df2 <- sizeRatings[sizeRatings$Word == 'ant' & sizeRatings$Language == 'notEnglish',]

df3 <- sizeRatings[sizeRatings$Class == 'plant',]
mean(df3$MeanFamiliarity)

df4 <- sizeRatings[sizeRatings$Class == 'animal',]
mean(df4$MeanFamiliarity)
```

Another useful operator is the `%in%` operator. This can be more efficient that many chained `or` statements.

```{r}
fruits <- sizeRatings[sizeRatings$Word %in% c('cherry','lemon','pear','blackberry'),]
```

We can also create new factors! In the code below, we add a new factor which has a value of `familiar` if `sizeRatings$MeanFamiliarity` is greater than 0.7, and `unfamiliar` if not. 

```{r}
sizeRatings$binary_MF <- ifelse(sizeRatings$MeanFamiliarity > 0.7, 'familiar', 'unfamiliar')
```

```
What differences are there between $Language and $binary_MF when you call str(sizeRatings) ?

Can you think of what differences there are between Factors and character vectors?
```
